---
title: "R Notebook"
output: html_notebook
---

## Create Wumpus World Matrix

Let's create a Matrix representing one Wumpus World.

- B |  Breeze
- S |  Smell
- G |  Glitter
- W |  Wumpus
- P |  Pit

B, S, and G can be combined on the same spot.
B and S occor only on non-diagonal fields in relation to the Wumpus or Pit
Glitter occurs only on a field with Gold
There is one Wumpus, 3 Pits and 1 Field with Gold

```{r}
w <- matrix(data="", nrow=4, ncol=4)
w
```

After creating the Matrix we can populate it. First let's position Wumpus Pit and Gold. This process we can randomize later.

```{r}

# TODO: Randomize!
w[3,1] = "W"
w[3,2] = "G"
w[1,3] = "P"
w[3,3] = "P"
w[4,4] = "P"
w

```

Now let's place the sensable cues: 

```{r}
# TODO: Abstract and create function with a Parameter to append the cue
fields <- which(w == "P", arr.ind=TRUE)
fields

for (i in 1:(length(fields)/2)) {
    row <- fields[[i, 1]]
    col <- fields[[i, 2]]
    
    top <- w[max(1, row-1),col]
    bottom <- w[min(4,row + 1), col]
    left <- w[row, max(1, col - 1)]
    right <- w[row, min(4, col + 1)]
      
    if (top != "W" | top != "P" |  top != "B") {
      w[max(1, row - 1), col] <- paste(top, "B", sep="")
    }
    if (bottom != "W" | bottom!= "P" | bottom!= "B") {
      w[min(4,row + 1), col] <- paste(bottom, "B", sep="")
    }
    if (left != "W" | left != "P" | left != "B") {
      w[row, max(1, col - 1)] <- paste(left, "B", sep="")
    }
    if (right != "W" | right != "P" | right != "B") {
      w[row, min(4,col + 1)] <- paste(right, "B", sep="")
    }
}
w

```



