---
title: "R Notebook"
output: html_notebook
---

## Create Wumpus World Matrix

Let's create a Matrix representing one Wumpus World.

- B |  Breeze
- S |  Smell
- G |  Glitter
- W |  Wumpus
- P |  Pit
- V |  Visited

B, S, and G can be combined on the same spot.
B and S occor only on non-diagonal fields in relation to the Wumpus or Pit
Glitter occurs only on a field with Gold
There is one Wumpus, 3 Pits and 1 Field with Gold

```{r}
w <- matrix(data="", nrow=4, ncol=4)
w
```

After creating the Matrix we can populate it. First let's position Wumpus Pit and Gold. This process we can randomize later.

w```{r}

# TODO: Randomize!
w[3,1] = "W"
w[3,2] = "G"
w[1,3] = "P"
w[3,3] = "P"
w[4,4] = "P"
w

```

Now let's place the sensable cues: 

```{r}

fields <- which(w == "P", arr.ind = TRUE)
fields

for (i in 1:(length(fields)/2)) {
    row <- fields[[i, 1]]
    col <- fields[[i, 2]]
    top <- w[max(1, row - 1),col]
    bottom <- w[min(4,row + 1), col]
    left <- w[row, max(1, col - 1)]
    right <- w[row, min(4, col + 1)]
      
    if (top != "W" | top != "P" |  top != "B") {
      w[max(1, row - 1), col] <- paste(top, "B", sep = "")
    }
    
    if (bottom != "W" | bottom != "P" | bottom != "B") {
      w[min(4,row + 1), col] <- paste(bottom, "B", sep = "")
    }
    
    if (left != "W" | left != "P" | left != "B") {
      w[row, max(1, col - 1)] <- paste(left, "B", sep = "")
    }
    
    if (right != "W" | right != "P" | right != "B") {
      w[row, min(4,col + 1)] <- paste(right, "B", sep = "")
    }
}
w

```




```{r}

# Agent State
agent <- c((row, column), "R", 1)

position <- agent[1]
orientation <- agent[2]
shots <- agent[3]

# Uncovered Gameboard
u <- w

while position != "G" {
  position <- agent[1]
  orientation <- agent[2]
  shots <- agent[3]

  u <- sense_environment(position) # Check if we have been here and if we need to sense at all?
  
  decide(u, position, orientation, shots) # Which action to take based on our KB? Move, rotate or shoot?
  
  orientation <- rotate() 
  position <- move()
  shots <- shoot()
  
  agent <- c(position, orientation, shots)
}
```



Sense the environment:

```{r}
sense_environment(row, column) <- function() {
  top <- w[max(1, row-1),col]
  bottom <- w[min(4,row + 1), col]
  left <- w[row, max(1, col - 1)]   
  right <- w[row, min(4, col + 1)]

  w[left:right, top:bottom]
}

```




When we sense from 1,1 we find one field to be unvisited with no senses coming from that: 2,2

```{r}
w[max(1,1-1):min(4,1+1),max(1, 4-1):min(4,4+1)]
```

Result:

     [,1] [,2]
[1,] "P"  "B" 
[2,] "B"  ""  



Traverse our Wumpus World by moving to safe spots based on sensing

```{r}

# Submatrix observable from 1,1:
sensed_submatrix = u[1:2, 1:2]
position = c(1,1)

# TODO: Only decide on empty fields after all Safe Fields (V, B) have been moved to and additional info is on the "u" map

# From submatrix get the coordinates from all fields, we did not touch yet, i.e. having empty string ""
fields_to_sense = which(sensed_submatrix == "", arr.ind = TRUE)

# Disregard the field we are on
for (i in 1:nrow(fields_to_sense)) { 
  row <- fields_to_sense[i, 1]
  col <- fields_to_sense[i, 2]
  
  maxcol <- ncol(sensed_submatrix)
  maxrow <- nrow(sensed_submatrix)
  
  top <- sensed_submatrix[max(1, row - 1),col]
  bottom <- sensed_submatrix[min(maxrow, row + 1), col]
  left <- sensed_submatrix[row, max(1, col - 1)]   
  right <- sensed_submatrix[row, min(maxcol, col + 1)]
  
 if (row == 1 & col == 1) {
  check(c(bottom, right))
 }
  
 if (row == 1 & col < maxcol) {
  check(c(bottom, left, right))
 }
 
 if (col == 1 & row < maxrow) {
  check(c(top, bottom, right))
 }
 
 if (row == maxrow & col < maxcol) {
  check(c(left, top, right))
 }
 
 if (col == maxcol & row < maxrow) {
  check(c(left, top, bottom))
 }
 
 if col == maxcol & row == maxrow {
  check(c(top, left))
 }
}

check(positions) <- function() {
  return !any(
    all(grepl("B", positions)),
    all(grepl("S", positions))
  )
}


```
