---
title: "RDF Introduction"
author: "Dominik"
date: "2024-06-21"
output: html_document
---

```{r setup, include=FALSE}

library(rdflib)
library(dplyr)
library(tidyr)
library(tibble)
library(jsonld)

```


## RDF Intro

In the world of data science, RDF is a bit of an ugly duckling. Like XML and Java, only without the massive-adoption-that-refuses-to-die part. In fact RDF is most frequently expressed in XML, and RDF tools are written in Java, which help give RDF has the aesthetics of steampunk, of some technology for some futuristic Semantic Web1 in a toolset that feels about as lightweight and modern as iron dreadnought.

But don’t let these appearances deceive you. RDF really is cool. If you’ve ever gotten carried away using tidyr::gather to make everything into one long table, you may have noticed you can just about always get things down to about three columns, as we see with an obligatory mtcars data example for tidyr::gather:

```{r}
mtcars
```


```{r cars}

car_triples <- mtcars %>% 
  rownames_to_column("Model") %>% 
  gather(attribute,measurement, -Model)

car_triples


```

If you like long tables like this, RDF is for you. This layout isn’t “Tidy Data,” where rows are observations and columns are variables, but it is damn useful sometimes. This format is very liquid, easy to reshape into other structures – so much so that tidyr::gather was originally known as melt in the reshape2 package. It’s also a good way to get started thinking about RDF.

## Subject URIs

Using row names as our subject was intuitive but actually a bit sloppy. tidyverse lovers know that tidyverse doesn’t like rownames, they aren’t tidy and have a way of causing trouble. Of course, we made rownames into a proper column to use gather, but we could have taken this one step further. In true tidyverse fashion, this rownames-column is really just one more variable we can observe, one more attribute of the thing we were describing: say, thing A (Car A) is a car_model_name as Mazda RX4 and thing A also has mpg of 21. We can accomplish such a greater level of abstraction by keeping the Model as just another variable the row ids themselves as the key (i.e. the subject) of our triple:

```{r}

car_triples <- 
mtcars %>% 
  rownames_to_column("Model") %>% 
  rowid_to_column("subject") %>% 
  gather(predicate, object, -subject)

car_triples

```

This is identical to a gather of all columns, where we have just made the original row ids an explicit column for reference (diligent reader will recognize we would need this information to reverse the operation and spread the data back into it’s wide form; without it, our transformation is lossy and irreversible). Our subject column now consists only of simple numeric id’s, while we have gained an additional triple for every row in the original data which states Model of each id number (e.g. 1 is Model Mazda RX4). Okay, now you’re probably thinking: “wait a minute, 1 is not a very unique or specific key, surely that will cause trouble,” and you’d be right. For instance, if we performed the same transformation on the iris data, we get triples in the exact same format, ready to bind_rows:

```{r}

iris_triples <- iris %>%
  rowid_to_column("subject") %>%
  gather(key = predicate, value = object, -subject)

iris_triples
```

```{r}

iris_triples <- iris %>%
  rowid_to_column("subject") %>%
  mutate(subject = paste0("http://example.com/", "iris#", subject)) %>%
  gather(key = predicate, value = object, -subject) %>%
  mutate(predicate = paste0("http://example.com/", "iris#", predicate)) %>%
  mutate(object = paste0("\"", object, "\"", "^^http://www.w3.org/2001/XMLSchema#double"))

iris_triples
```

```{r}

rdf <- rdf()

base <- paste0("http://www.semanticweb.org/dominikb1888/ontologies/2024/4/", "Pizza#")


rdf %>%

  rdf_add(
    subject = paste0(base, "Example-Margerita2"),
    predicate = paste0(base, "hasCalorificValue"),
    object = 293
  )

options(rdf_print_format = "turtle")

rdf

```


```{r}



example_pizza <- list("Example-Margerita2", "hasCalorificValue", 293 )
base <- paste0("http://www.semanticweb.org/dominikb1888/ontologies/2024/4/", "Pizza#")

rdf <- as_rdf(
    example_pizza,
    base = base,
    key_column = 1,
  ).list(R, context)

options(rdf_print_format = "turtle")

rdf



```

